\documentstyle[epsfig,amsmath,11pt]{article}
%\usepackage{amssymb,psfig,epsfig}
\bibliographystyle{plain}
\setlength{\oddsidemargin}{0.01in}
\setlength{\textwidth}{6.8in}
\setlength{\topmargin}{-0.51in}
\setlength{\textheight}{9.5in}

\begin{document}
\def\bebf{\mbox{\boldmath $\beta$}}
\def\mubf{\mbox{\boldmath $\mu$}}
\def\epbf{\mbox{\scriptsize\boldmath $\epsilon$}}
\def\mus{\mbox{\scriptsize\boldmath $\mu$}}


\title{BIOS 622 Homework 6}
\vspace{0.2in}
\author{Instructor: Sijin Wen\\ }
\maketitle

\baselineskip=14pt


 

\begin{enumerate}


\item The data of time to death ($T_1,...,T_n$) are given as follows:

\begin{itemize}
\item Steroid: 1, 1+(3), 4+, 5, 7, 8, 10, 10+, 12+, 16+(3)
\item Control: 1+, 2(3), 3(2), 3+, 5+(2), 16+(6)
\end{itemize}

\begin{enumerate}

\item This homework is based on Weibull distribution assumption as follows. Let $Z=1$ for the steroid and $Z = 0$ for the control. Assume that survival time $T|Z$ is Weibull with scale parameter $\lambda_Z$ and shape parameter $\gamma$ (exponential if $\gamma = 1$) and the survival function is 
$$
S_Z(t) = exp\{-\lambda_Z t^{\gamma}\}
$$
where
$$
\lambda_Z = \frac{1}{exp\{(\beta_0 + \beta_1 Z)\gamma\}} = exp\{(-\beta_0 - \beta_1 Z)\gamma\}
$$


Write down the corresponding hazard function $h_Z(t)$

\item Write down the likelihood function on $\beta_0$, $\beta_1$ and $\gamma$;  Explain briefly how you might find the maximum-likelihood estimates of $\beta_0$, $\beta_1$ and $\gamma$ (do not attempt to find them)

\item Find estimates of $\beta_0$, $\beta_1$ and $\gamma$ using proc 'lifereg' in SAS (or 'survreg' in R). What is the p-value on $\beta_1$? Please state the null and alternative hypotheses. Is there a big difference on p-value, in comparison with log-rank test in (a)?

\item What is the predicted median survival time for each group under the Weibull assumption?

\item What is hazard ratio of treatment effect and its 95\% CI (you may use the R package/library `SurvRegCensCov') ? 

\item A survival model is considered as an accelerated failure time (AFT) model if $S(t | Z=1) = S(t/\phi | Z = 0)$ where $\phi^{-1}$ is accelerated factor (a constant). Is this Weibull model a AFT? If yes, what is the value of the accelerated factor $\phi^{-1}$ (hint: review course notes Page 137, 143 and 146). 

\item Estimate the median follow-up time for each group using reverse Kaplan-Meier estimation.

\end{enumerate}

\item  (Bonus) Let $S(t)$ is the survival function of $T$, show that
$$
E(T) = \int S(t) dt.
$$ 
This implies that the mean survival time is the area under the survival function/curve (relating to the bonus question in homework 5: a non-parametric way to find mean survival time).


\end{enumerate}






\end{document}





