\documentstyle[epsfig,amsmath,11pt]{article}
%\usepackage{amssymb,psfig,epsfig}
\bibliographystyle{plain}
\setlength{\oddsidemargin}{0.01in}
\setlength{\textwidth}{6.8in}
\setlength{\topmargin}{-0.51in}
\setlength{\textheight}{9.5in}

\begin{document}
\def\bebf{\mbox{\boldmath $\beta$}}
\def\mubf{\mbox{\boldmath $\mu$}}
\def\epbf{\mbox{\scriptsize\boldmath $\epsilon$}}
\def\mus{\mbox{\scriptsize\boldmath $\mu$}}


\title{BIOS 622 Midterm Exam}
\vspace{0.2in}
\author{Instructor: Sijin Wen\\ }
\date{March 20, 2025}
\maketitle

\baselineskip=14pt

 

\begin{enumerate}


\item Consider a study carried out to evaluate the effect of a drug for high blood pressure. Elderly volunteers are randomly allocated to one of three treatments: a placebo, the drug that is currently most common, and the new drug that has been developed. The study lasts for ten years, and the event of interest is occurrence of any kind of heart disease, including death resulting from heart disease. Briefly describe a situation from this study that might lead to an observation that is right-censored.

 \vspace{7 cm}

\item The lifetime of light bulbs follows an exponential distribution with hazard rate of 0.001 failures per hour of use. (a) Find the mean lifetime of a randomly selected light bulb. (b) Find the median lifetime of a randomly selected light bulb. (c) What is the probability of a light bulb will still function after 2000 hours of use?  

\vspace{7 cm}

\item A survival time $T$ has its hazard function as $h(t) = 2t$. (1) Which parametric statistical distribution does the survival time follow (such as Exponental, Weibull, log-normal, etc.)? (2) Find the survival function and density function of $T$. 

\vspace{7 cm}


\item Let $X$ be a random variable with $E(X) =\mu$ (mean) \& $Var(X) = \sigma^2$ (variance). 
\begin{enumerate}
\item Use Delta method to find mean and variance of $Y = g(X)$ \\(Hint: let $Y = g(X) = g(\mu) + (X - \mu) g'(\mu)$ based on Taylar expansion). 

\item In a special case when $g$ takes a natural log-transformation, that is $Y = g(X) = ln(X)$, find $E(Y)$ and $Var(Y)$, assuming $\mu$ and $\sigma^2$ are given (they can be estimated from historical data, for example). 

\item Recall $H(t) = - ln(S(t))$ in the course slides, please find a formula for the estimate of $\hat{Var(H(t))}$, assuming $\hat{Var(S(t))}$ is already obtained from Greenwood method (say).

\item Redo (b) and (c) for $Y= \exp(X)$.
 
\end{enumerate}


\vspace{14 cm}

\item A modified example from the course notes (Page 60): the data of time to death ($t_i$) from severe viral disease are given as follows:

\begin{itemize}
\item Steroid: 1(3), 1+, 4+, 5, 7, 8, 10, 10+, 12+, 16+(3)
\item Control: 1+, 2, 3(2), 3+, 5(2), 16+(8)
\end{itemize}

Define:

$D_i$ = Number of events (deaths) at time $t_i$

$S_i$ = Number that are known to have survived beyond $t_i$ (by convention: includes those censored at $t_i$)

$N_i$ = Number of `at risk' of being observed to die at time $t_i$, that is $N_i= S_i + D_i$

$h_i(t)$ = hazard function at time $t_i$

$H_i(t)$ = corresponding cumulative hazard

\begin{enumerate}
\item For each treatment group, complete the following table with Kaplan-Meier estimator of survival function:

%\begin{table}
%\caption{}
\begin{center}

\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
 \hline
\hline
 $i$ & $t_i$ & $N_i$  & $D_i$ & $S_i$ & $\hat{S_i (t)}$ & s.e.$(\hat{S_i(t)})$ & $\hat{H_i(t)}$ & s.e.($\hat{H_i(t)}$) & $\hat{h_i(t)}$ 
\\
\hline
1 & & & & & & & & &\\
\hline
2 & & & & & & & & &\\
\hline
.& .& . & . & .& .&. &. &. &.\\
\hline
k & & & & & & & & &\\
\hline
\hline

\end{tabular}
\end{center}
%\end{table}

\vspace{1cm}

\item Find 95\% confidence interval for log($S_i (t)$) at each time point in Table (a) (hint: you may use the result from Q4(b)) .
\item Find 95\% confidence interval for $S_i (t)$ at each time point in Table (a) using the exponential transformation from (b).
\item For the Steroid group, assume that survival times are distributed according to a Weibull distribution with parameters $\lambda$ and $\gamma$, the hazard and survival functions can be written as
$$
h(t) = \lambda \gamma t^{\gamma -1}
$$
$$
S(t) = e^{-\lambda t^{\gamma}}
$$

(i) Write down the likelihood function for the Steroid data; (ii) Explain briefly how you might find the maximum-likelihood estimates of $\lambda$ and $\gamma$ (do not attempt to find them); (iii) (Bonus) find the maximum-likelihood estimates of $\lambda$ and $\gamma$.
\item Plot Kaplan-Meier survival curve (using either R or SAS) for each group. Can you assume the survival time follows Weibull distribution and why?   
\item Find a p-value from log-rank test between two groups by hand (by yourself). Double-check the result with SAS or R function. Please state the null and alternative hypotheses.
\item Estimate median survival time and its 95\% confidence interval for each group from SAS or R function first, and do it by hand (by yourself) late.
\item Estimate the difference of median survivals (Control - Steroid) between two groups and its 95\% confidence interval (hint: you can make reasonable assumptions (parametric model or non-parametric model), and use all methods you have learnt such as MLE, Delta method, Bayesian method, bootstraping,..., etc). 

\end{enumerate}




\end{enumerate}



\end{document}





